---
title: "Study_hours dataset"
output: html_notebook
---

# Install and load necessary libraries
```{r}
library(ggplot2)
```

# Load the dataset
```{r}
study_hours =read.csv("E://1st Sem/PROGRAMMING/Study hours/study hours .csv")
```

# View the first few rows
```{r}
head(study_hours,5)
```
#Summary
```{r}
summary(study_hours)
```
# Inspect the data
```{r}
str(study_hours)
```
# Clean the data by handling missing values
```{r}
is.na(study_hours)
```
#Finding null values
```{r}
colSums(is.na(study_hours))
```
# Data Visualization - Scatter plot
```{r}
ggplot(study_hours, aes(x = Hours, y = Scores)) +
  geom_point(color = "purple") +
  ggtitle("Scatter Plot of Study Hours vs. Scores") +
  xlab("Study Hours") +
  ylab("Scores")
```

# Find unique values in the "Hours" column
```{r}
unique(study_hours$Hours)

unique(study_hours$Scores)
```
# Box plot 
```{r}
ggplot(study_hours, aes(x = Hours, y = Scores)) +
  geom_boxplot() +
  ggtitle("Scatter plot of hours vs. scores")
```

# Build the regression model predicting Scores based on Hours
```{r}
model <- lm(Scores ~ Hours, data = study_hours)
summary(model)
```

# Create a scatter plot with the regression line
```{r}
ggplot(study_hours, aes(x = Hours, y = Scores)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  ggtitle("Scatter plot of Hours vs. Scores with Regression Line")
```
# Create a histogram of study hours
```{r}
ggplot(study_hours, aes(x = Hours)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  ggtitle("Histogram of Study Hours") +
  xlab("Hours") +
  ylab("Frequency")
```

# Create a histogram of scores
```{r}
ggplot(study_hours, aes(x = Scores)) +
  geom_histogram(binwidth = 5, fill = "green", color = "black") +
  ggtitle("Histogram of Scores") +
  xlab("Scores") +
  ylab("Frequency")
```
```{r}
predictions <- predict(model, newdata = study_hours)
```


# Compare the predictions with the actual values
```{r}
dim(Hours)

# Check the length of predictions
length(predictions)

# Create a new dataframe with predictors (assuming 'new_study' is your new data)
new_study <- data.frame(Hours = c(1, 2, 3))  # Example: Input new study hours

# Predict using the model with new data
new_predictions <- predict(model, newdata = new_study)
new_predictions

```

# Visualize the distribution of study hours]
```{r}
ggplot(study_hours, aes(x = Hours)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Study Hours", x = "Study Hours per Day", y = "Frequency")
```

ggplot(data, aes(x = StudyHours)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Study Hours", x = "Study Hours per Day", y = "Frequency")

# Calculate the correlation between study hours and performance scores
correlation <- cor(data$StudyHours, data$PerformanceScore)
print(paste("Correlation between Study Hours and Performance Scores: ", correlation))

# Visualize the correlation
```{r}
ggplot(study_hours, aes(x = Hours, y = Scores)) +
  geom_point(color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Correlation between Study Hours and Performance Scores",
       x = "Study Hours per Day", y = "Performance Score")

```


